**

### 2010-05-20 BD_disc/old.tex
minor typo change
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2278 25464f21-6099-402c-b30a-640f1d40d38b

Some studies of disc lifetimes find no significant difference between
BD stars and their higher mass counterparts
\citep{2003AJ....126.1515J}, however a possible correlation of disc
lifetime with stellar mass is proposed in \cite{2007ApJ...660.1517S}.
Additionally, \cite{2009arXiv0901.2603K} find that, for TTS older than
3 Myrs, the disc dispersal time is mass dependent for intermediate
mass (IM) to solar-type stars. This result only holds for the mass
transition stated (IM to solar-type) and is based on stars with
combined dust signatures, from infrared (IR) excesses and accretion
indicators (using a EW[H$\alpha$], criterion).
\cite{2009arXiv0901.4120D} also find, for TTS, a mass dependent
general SED shape for disc candidates, with higher mass stars showing
a featureless continuum at $\lambda>$8.0$\mu$m with lower mass stars
generally showing strong silicate and PAH features at 10 and 2$\mu$m,
indicative of mass dependent disc evolution. Whether in fact the disc
lifetimes and therefore rate or type of involved physical processes
are a function of stellar mass could provide another piece of evidence
separating the previously assumed continuous (in mass) evolutionary
principles of stars into a dichotomy separated by the hydrogen-burning
mass limit, and alter the probability of forming planets in the BD
regime.

@@

Several mechanisms may act to completely dissipate a circumsteller
disc. These include accretion from the disc ($\delta M_{\rm
  disc}=\dot{M}$), photoevaporation or photoionisation due to flux
from the central and neighbouring stars, and the formation of giant
planets or interaction with binary partners. However, the entire disc
does not have to be dissipated to be missed by disc fraction surveys.
More subtly, disc fractions can be altered due to wavelength dependent
detection. This is due to the peak flux of the thermal blackbody
emission of the disc being proportional to the temperature,
$\lambda^{\rm thermal}_{\rm peak}\propto T_{\rm disc}$, and the
temperature in the disc being approximately a function of distance
from the star. Thus for a given system, as one increases the
wavelength of observation, the corresponding radial position of the
effective source will increase. Moreover, as the disc flux must be
separated from the photospheric flux the important quantity is the
separation in wavelength between the emission peaks for the stellar
and thermal disc components. Therefore the presence of an inner disc
hole can change derived disc fractions, as a function of
wavelength. Interestingly, \cite{2009arXiv0901.2603K} find that
results for disc dissipation timescales based only on their accretion
indicators or inferred disc presence do not agree, perhaps due to
inclusion of transitional objects, where the inner disc is being
cleared, and is therefore not detected at their target
wavelength. \cite{2009arXiv0901.2603K} suggest that this is caused by
photoevaporation of the inner disc, leading to a mass dependent disc
lifetime, at their wavelength, for their sample. However, as discussed
in \cite{2009arXiv0901.4120D}, the inner hole can be created via dust
grain growth and midplane settling, giant planet formation or
photoevaporation by ultraviolet (UV) and far-UV (FUV) flux
\citep{2007MNRAS.378..369N}.  \cite{2009arXiv0901.4120D} argue that if
binarity or giant planet formation or photoevaporation cause inner
disc clearing the resulting system would not show active accretion
from radii below the inner disc boundary. For the case of clearing by
dust grain growth and settling accretion from a hot gas disc within
the inner radius is still possible, and indeed evidence for such a hot
gas inner disc has been found through SED fitting
\citep{2005ApJ...623..952E,2008ApJ...689..513T}. Additionally,
\cite{2009arXiv0901.4445E} argue that for M stars, with a pure
reprocessing disc where the inner disc radius ($R_{\rm inner}$) is
equal to the dust sublimation radius ($R_{\rm dust}$, discussed in
Section \ref{physics}), the stellar and thermal components are not
separable at wavelengths shorter than 6$\mu$m. In the case of
\cite{2009arXiv0901.4445E}, the inner disc hole was created by
photoionising radiation from the central star destroying the dust
component of the disc. Recent direct evidence for a photoevaporating
disc has been found for the circumstellar disc around a low mass TTS
in the Sigma Orionis cluster \citep{2009arXiv0902.0457R}. The
detection was characterised by a tenuous photoevaporation wind and
little or no accretion, and a disc detection at 24$\mu$m indicative of
a large inner hole. Additionally, for BDs, \cite{2009arXiv0902.0638M}
find that the disc mass and radius distributions in the Orion
Trapezium cluster are consistent with photoevaporation over 1 Myr (the
adopted age of the cluster), from UV and FUV flux generated by the
massive $\theta^1$ Ori C.

@@

Although the creation of the inner disc hole could have several
causes, at least part of the effect is undoubtedly due to
accretion. Accretion acts to directly deplete the inner disc (where
the material is accreted from). Accretion also depletes the inner disc
indirectly, as accretion generates hot continuum emission
\citep{1998ASPC..154.1709G}, increasing the photoionising UV and FUV
flux available \citep{1998ASPC..154.1709G}, and therefore affecting
the derived disc fraction \citep{2009arXiv0901.3684H}. This effect may
be amplified for BDD systems where the combined photospheric emission
is dominated by accretion flux. However, accretion columns modeled by
\cite{2004MNRAS.348..879A} were found to produce negligible levels of
photoionising flux due in part to the high optical depth of the column
itself to Lyman continuum photons. Additionally, due to the high
concentrations of hydrogen in the accretion column and hotspot the
emission is more likely to be characterised by a stellar atmosphere
than a black body. However, for our work as emission from a stellar
atmosphere or a black body of equal temperature are approximately
equal for wavelengths longer than the Lyman edge we have neglected
this effect. Whilst the accretion column and hotspot may not
contribute to the photoinising flux within the system they will of
course alter the disc evolution by an increase of the emitted
flux. This increase in flux will lead to an increase in temperature of
the disc inner and outer regions, and more particulary a change in the
dust sublimation effects at the inner edge.

The current proposed mechanism of accretion is a magnetically
channelled accretion `column', where material from a truncated (at the
co-rotation radius $R_{\rm co}$) inner disc boundary falls onto the
surface of the star along magnetic field lines
\citep{1990RvMA....3..234C,1991ApJ...370L..39K}. This mechanism is
widely accepted, with strong evidence found from the excessive widths
of H$\alpha$ ($>$10$\AA$) emission lines, over that caused by
chromospheric activity and stellar rotation
\citep{1998ASPC..154.1793W}, caused by the high velocities of
infalling matter, ~100-200 kms$^{-1}$ \citep{1998ApJ...492..743M}.

@@

If disc-locking is present for BDD systems, as for TTS stars, one
would expect accretion of material from a magnetic truncation radius,
$R_{\rm T}$ \citep{2007prpl.conf..479B}.  \cite{2007prpl.conf..479B}
show that the magnetic truncation radius, $R_{\rm T}$, must be less
than the co-rotation radius, $R_{\rm co}$ (where the Keplerian angular
rotation rate of the disc is the same as the angular rotation rate of
the star), as material outside this radius will have too much angular
momentum to fall onto the star. This is implicitly adopted in the
proposed accretion mechanism, where the disc is magnetically
truncated, and therefore locked to the star. This magnetic truncation
would lead to an inner wall (and inner hole) of the accretion
disc. Evidence for an inner disc wall has been found from near-IR
excesses from the inner rims of Herbig Ae stars, with emission peaking
at 2-3$\mu$m \citep{2001ApJ...560..957D}. The actual shape and radial
position of this wall is affected by several factors, discussed
further in Sections \ref{dust_edge} and \ref{inner_edge}. The previous
work of \cite{1997AJ....114..288M} explored the effects of simple mass
accretion and the radius to the inner disc on the infrared colours of
CTTS, useing a prescribed range of inner disc
radii. \cite{1997AJ....114..288M} find that the scatter in infrared
colours for there parameter range, $R_{\rm inner}=$1$-$12$R_*$ and
$\dot{M_{\rm acc}}=$10$^{-9}$to 10$^{-5} M_{\odot}yr^{-1}$, matches
the scatter of the CTTS locus found in the data. The models of
\cite{1997AJ....114..288M} used physically flat discs however meaning
their results will not stand for discs with significant flaring. More
importantly for our work \cite{1997AJ....114..288M} find a correlation
of the infrared locus of models with the accretion and inner disc
location. Essentially, changes in the stellar flux levels and position
of the inner wall lead to temperature changes of this inner edge and
therefore changes in the balance of flux across the infrared
wavelength regime. As this study does not include the effects of
vertical hydrostatic equilibirum or even the study of flared discs one
would expect, perhaps, a more complex relationship between the inner
edge position and infrared excess in `real' systems. This means that
studies aiming to provide evidence of disc locking by comparing disc
indicators with rotation rates may include unforseen biases. As the
rotation rate will effect the co-rotation radius and therefore
truncate the disc at different radii, thereby enhancing or suppressing
any relationship of infrared excess (or disc presence) with rotation
rate.

In summary, several key areas in the theoretical studies of BDs hinge
critically on observational constraints, where the derived values are
heavily affected by disc and/or accretion presence.  Formation
mechanisms, IMFs, disc dissipation timescales, accretion rate--stellar
mass relationships and studies of disc-locking mechanisms using
correlations of rotation rates with infrared excesses, all require
derivations of either masses, ages or disc fractions (or a
combination), supplemented by other parameters. The consequences of
the disc structure for BD accretion discs, particularly vertical
structure, given the assumption of vertical hydrostatic equilibrium,
have already been discussed for a limited range of BD parameters in
\cite{2004MNRAS.351..607W}. Indeed \cite{2004MNRAS.351..607W} have
shown that BDD systems seen at higher inclinations can be confused
with CTTS systems, having similar colours and
magnitudes. Additionally, a benchmark model grid of spectra for star
and disc systems in the mass range 0.1$M_{\odot}<M_*<$50$M_{\odot}$
has previously been published in \cite{2006ApJS..167..256R}.

@@

In this study we assume that accretion from the central star occurs
along magnetically channelled columns from the inner disc boundary.
For TTS stars, \cite{2007prpl.conf..479B} show that the magnetic
truncation radius ($R_{\rm mag}$) is less than the co-rotation radius
($R_{\rm co}$), where the angular Keplerian velocity of the
disc is equal to the surface angular velocity of the central star.
Calculations of the magnetic truncation radius depend on derivations
of the surface magnetic field \citep{1991ApJ...370L..39K}. This is
currently unavailable for BD stars, due to increased molecular species
obscuring the Zeeman splitting signatures, that are normally used to
derive stellar surface magnetic fields. Therefore, for our model grid
we have adopted an initial inner disc radius as the co-rotation
radius,
\begin{equation}
{R_{\rm inner}=({GM_*\tau^2\over{4\pi ^2}})^{1\over{3}}},
\label{inner_eq}
\end{equation}
 where $\tau$ is the stellar rotation period and $R_{\rm inner}$ is
the inner radius. This is effectively adopting a disc-locking
mechanism (without associated angular momentum loss), as for
disc-locked stars, $R_{\rm mag}\approx R_{\rm co}$,
\citep{1991ApJ...370L..39K,1994ApJ...429..781S}. Therefore, for our
model simulations, this inner edge radius is dependent on, and derived
from, the adopted value of the rotational period for the central star,
as well as being weakly dependent on the stellar mass. As discussed in
Section \ref{intro}, the inner disc can be cleared through a number of
mechanisms, including binarity or Giant planet formation,
photoevaporation or photoionisation of the disc and dust grain growth
or settling. For BDD systems where a disc is modeled a treatment of
dust sublimation due to photoionising UV and FUV flux is included
(discussed in Section \ref{dust_edge}). However, the effects of
binarity or Giant planet formation are neglected. Further to the
stellar mass and period required prior to calculation of the inner
disc radius, we require a disc mass (in stellar masses).

@@

\begin{figure}
  \vspace*{174pt}
  \special{psfile="Fig/albedo.ps"
   hoffset=235 voffset=0 hscale=30 vscale=30 angle=90}
 \caption{Figures of the albedo (\textit{top panel}), and scattering
   (dashed line) and absorption (solid line) opacties (\textit{bottom
     panel}) against log($\lambda$) (in $\mu$m) for our adopted dust
   population. For \textit{both panels} the vertical dashed line is
   plotted at 10$\mu$m to higlight the silicate features.
   \label{albedo}}
\end{figure}

@@

The practical implementation is as follows: An initial coarse
approximate temperature structure is found for the grid without dust,
only a fixed gas distribution. Then, first three full \textit{Lucy}
algorithm iterations are run for a grid with a maximum cell opacity of
0.1 (with dust added to this level). Three more iterations are then
performed with dust opacity limited to $\tau=$1 per cell. A further
three \textit{Lucy} algorithm iterations are then performed with the
full dust opacity, where dust is added until the gas-to-dust ratio is
achieved($\epsilon =$100). For each iteration (from the first to the
ninth) dust in cells which reach the sublimation temperature and
density is removed from the disc by setting the opacity to zero for
the cell. Once these nine iterations are complete, the grid is refined
around the disc photosphere, and the radiative transfer alogrithm is
iterated through to convergence. After this first convergence, a
vertical hydrostatic equilibrium algorithm is run, after which the
disc grid is unrefined and then re-refined around the disc
photosphere. This process continues for five hydrostatic equilibrium
calculations. Finally, once the hydrostatic equilibrium iteratins are
complete we perform a final dust sublimation run to remove any cells
where the dust temperature has moved above the sublimation temperature
and further refine the photosphere and inner edge of the disc (whilst
also unrefining cells where the opacity has reduced significantly and
high resolution is no longer required). In this model the dust
sublimation is effectively digital i.e. when the density and
temperature for sublimation is reached for a given cell the dust is
destroyed. We did attempt a more realistic model where the dust
sublimation causes the dust to be destroyed at an evanescently
decaying rate into the particular region. This was expected to lead to
a more realistic situation where the dust destruction propogates into
the disc material, and does not instantly destroy an entire
cell. However, this implementation lead to incremental sections of
dust being removed after each iteration, resulting in little change in
the overall density structure of the disc. This is perhaps interesting
for real systems, in that the dust sublimation should not be expected
to act as a digital `switch'.

@@

Binarity and Giant planet formation would be expected to clear an
inner disc and shut down accretion from within the inner wall due to
gravitational perturbations. In addition, photoevaporation from UV and
UV flux is expected to halt accretion from radii within the inner disc
due to the removal of this material via a stellar/disc wind
\citep{2006MNRAS.369..229A}. These two clearing mechanisms could be
differentiate by the resulting masses of the discs
\citep{2007MNRAS.378..369N}.  However, for our model systems we have
neglected both of these effects.

@@

Once the radiative transfer code was completed simulated SEDs were
generated. These SEDs can be generated for any distance and for any
system-observer inclination. For our models we have set the distance,
again, to 10pc to create absolute flux SEDs and selected ten
inclinations (0, 27, 39, 48, 56, 64, 71, 77, 84 and 90$^{\circ}$). A
further useful feature of the {\sc torus} code is that the emitted
photons which make up the SED are tagged on their way to the
observer. These tags separate the photons into four groups. Firstly,
photons are separated by source into thermal (disc) or stellar groups.
These groups are then subdivided into photons which reach the observer
either directly or after scattering (for example see Figure
\ref{disc_eg_173}). The resulting SEDs are discussed in Section
\ref{seds}.

@@

This section details the range of each of the input parameters we have
varied and, where possible, gives justification for the selected
ranges from published observations. The simulations in this paper
cover variations in the stellar mass, stellar age, stellar rotation
rate, accretion rate, the areal coverage of the accretion stream, disc
mass fraction and the system inclination. In a subsequent paper we
introduce variation of the outer radius and using an analytical
vertical disc structure as oppose to vertical hydrostatic equilibrium.
.A summary of the values adopted for each input variable is shown in
Table \ref{par_space_table}.

@@

Typical disc lifetimes for solar type stars are of order $\approx$ 10
Myrs \citep{2001ApJ...553L.153H}. Therefore, we have adopted input
ages of 1 and 10 Myrs for our model grid, to span the approximate
range of ages over which discs influence will be important.

@@

Little observational evidence can be found for approximate sizes of
accretion hot spots due to their more transient nature and often
smaller coverages, when compared to cooler or `plage' spots
\citep{2007prpl.conf..297H}. \cite{1995A&A...299...89B} modeled the
size of the cool spots on solar-type stars for a selection of
periodically variable candidates. They found projections of cooler
spots, onto the stellar disc, of a few to ~60\%.
\cite{1995A&A...299...89B} also found projected sizes, onto the
stellar disc, of typically a few \% to around 10\% for hot spots.
\cite{1996AJ....112.2159B} used observations of YY Orionis monitoring
flux amplitude variations as a function of wavelength to derive a
probable hot spot area of around 10\%. The spot temperature was also
modeled for YY Orionis in \cite{1996AJ....112.2159B}, resulting in a
best fitting areal coverage of 11\%. Therefore, to bound the probable
areal coverage range of the accretion hot spots we have adopted areal
coverages of 1 and 10\%.

@@

GOT HERE NEEDS ALOT OF RE-READING FROM HERE ON IN**

@@

For our model grid we have selected a range of input parameters
covering expected, from observations, ranges. We have input the
fundamental parameters of age and mass by adoption of a set of
theoretical isochrones. We have intrinsically included a correlation
of rotation rate with initial inner edge position by setting this at
the co-rotation radius, which is expected to result in a correlation
of rotation rate with IR excess \citep{1997AJ....114..288M} (the
reverese of that expected due to angular momentum loss through
disc-locking). We have not assumed a relationship between accretion
rate and stellar mass for our model grid. Therefore, analysis of our
model grid in an attempt to recover the fundamental parameters of age
and mass, and show accretion rate to stellar mass or rotation rate to
infrared excess relationships, will allow us to assess the reliability
of these current theoretical models or relationships.

>>>>

Some studies of disc lifetimes find no significant difference between
BD stars and their higher mass counterparts
\citep{2003AJ....126.1515J}, however a possible correlation of disc
lifetime with stellar mass is proposed in \cite{2007ApJ...660.1517S}.
Additionally, \cite{2009arXiv0901.2603K} find that, for TTS older than
3 Myrs, the disc dispersal time is mass dependent for intermediate
mass (IM) to solar-type stars. This result only holds for the mass
transition stated (IM to solar-type) and is based on stars with
combined dust signatures, from infrared (IR) excesses and accretion
indicators (using a EW[H$\alpha$], criterion).
\cite{2009arXiv0901.4120D} also find, for TTS, a mass dependent
general SED shape for disc candidates, with higher mass stars showing
a featureless continuum at $\lambda>$8.0$\mu$m and lower mass stars
generally showing strong silicate and PAH features at 10 and 2$\mu$m,
indicative of mass dependent disc evolution. If the disc lifetimes and
therefore rate or type of involved physical processes are a function
of stellar mass. This could provide another piece of evidence
separating the previously assumed continuous (in mass) evolutionary
principles of stars. Resulting in a dichotomy separated by the
hydrogen-burning mass limit. This would also lead to a change in the
probability of forming planets in the BD regime, compared to CTTS.

@@

Several mechanisms may act to completely dissipate a circumsteller
disc. These include accretion from the disc ($\delta M_{\rm
  disc}=\dot{M}$), photoevaporation or photoionisation due to flux
from the central and neighbouring stars, and the formation of giant
planets or interaction with binary partners. However, the entire disc
does not have to be dissipated to be missed by disc fraction surveys.
More subtly, disc fractions can be altered due to wavelength dependent
detection. This is due to the peak flux of the thermal blackbody
emission of the disc being proportional to the temperature,
$\lambda^{\rm thermal}_{\rm peak}\propto T_{\rm disc}$, and the
temperature in the disc being approximately a function of distance
from the star. Thus for a given system, as one increases the
wavelength of observation, the corresponding radial position of the
effective source will increase. Moreover, as the disc flux must be
separated from the photospheric flux the important quantity is the
separation in wavelength between the emission peaks for the stellar
and thermal disc components. Therefore the presence of an inner disc
hole can change derived disc fractions, as a function of
wavelength. Interestingly, \cite{2009arXiv0901.2603K} find that
results for disc dissipation timescales based only on their accretion
indicators or inferred disc presence do not agree, perhaps due to
inclusion of transitional objects, where the inner disc is being
cleared, and is therefore not detected at their target
wavelength. \cite{2009arXiv0901.2603K} suggest that this is caused by
photoevaporation of the inner disc, leading to a mass dependent disc
lifetime, at their wavelength, for their sample. However, as discussed
in \cite{2009arXiv0901.4120D}, the inner hole can be created via dust
grain growth and mid plane settling, giant planet formation or
photoevaporation by ultraviolet (UV) and far-UV (FUV) flux
\citep{2007MNRAS.378..369N}.  \cite{2009arXiv0901.4120D} argue that if
binarity or giant planet formation or photoevaporation cause inner
disc clearing the resulting system would not show active accretion
from radii below the inner disc boundary. For the case of clearing by
dust grain growth and settling accretion from a hot gas disc within
the inner radius is still possible, and indeed evidence for such a hot
gas inner disc has been found through SED fitting
\citep{2005ApJ...623..952E,2008ApJ...689..513T}. Additionally,
\cite{2009arXiv0901.4445E} argue that for M stars, with a pure
reprocessing disc where the inner disc radius ($R_{\rm inner}$) is
equal to the dust sublimation radius ($R_{\rm dust}$, discussed in
Section \ref{physics}), the stellar and thermal components are not
separable at wavelengths shorter than 6$\mu$m. In the case of
\cite{2009arXiv0901.4445E}, the inner disc hole was created by
photoionising radiation from the central star destroying the dust
component of the disc. Recent direct evidence for a photoevaporating
disc has been found for the circumstellar disc around a low mass TTS
in the Sigma Orionis cluster \citep{2009arXiv0902.0457R}. The
detection was characterised by a tenuous photoevaporation wind and
little or no accretion, and a disc detection at 24$\mu$m indicative of
a large inner hole. Additionally, for BDs, \cite{2009arXiv0902.0638M}
find that the disc mass and radius distributions in the Orion
Trapezium cluster are consistent with photoevaporation over 1 Myr (the
adopted age of the cluster), from UV and FUV flux generated by the
massive $\theta^1$ Ori C.

@@

Although the creation of the inner disc hole could have several
causes, at least part of the effect is undoubtedly due to
accretion. Accretion acts to directly deplete the inner disc (where
the material is accreted from). Accretion also depletes the inner disc
indirectly, as accretion generates hot continuum emission
\citep{1998ASPC..154.1709G}, increasing the photoionising UV and FUV
flux available \citep{1998ASPC..154.1709G}, and therefore affecting
the derived disc fraction \citep{2009arXiv0901.3684H}. This effect may
be amplified for BDD systems where the combined photospheric emission
is dominated by accretion flux. However, accretion columns modeled by
\cite{2004MNRAS.348..879A} were found to produce negligible levels of
photoionising flux due in part to the high optical depth of the column
itself to Lyman continuum photons. Additionally, due to the high
concentrations of hydrogen in the accretion column and hotspot the
emission is more likely to be characterised by a stellar atmosphere
than a black body. However, for our work as emission from a stellar
atmosphere or a black body of equal temperature are approximately
equal for wavelengths longer than the Lyman edge we have neglected
this effect. Whilst the accretion column and hotspot may not
contribute to the photoinising flux within the system they will of
course alter the disc evolution by an increase of the emitted
flux. This increase in flux will lead to an increase in temperature of
the disc inner and outer regions, and more particularly a change in the
dust sublimation effects at the inner edge.

The current proposed mechanism of accretion is a magnetically
channelled accretion `column', where material from a truncated (at the
co-rotation radius $R_{\rm co}$) inner disc boundary falls onto the
surface of the star along magnetic field lines
\citep{1990RvMA....3..234C,1991ApJ...370L..39K}. This mechanism is
widely accepted, with strong evidence found from the excessive widths
of H$\alpha$ ($>$10$\AA$) emission lines, over that caused by
chromospheric activity and stellar rotation
\citep{1998ASPC..154.1793W}, caused by the high velocities of
infalling matter, $~$100-200 kms$^{-1}$ \citep{1998ApJ...492..743M}.

@@

If disc-locking is present for BDD systems, as for TTS stars, one
would expect accretion of material from a magnetic truncation radius,
$R_{\rm T}$ \citep{2007prpl.conf..479B}. \cite{2007prpl.conf..479B}
show that the magnetic truncation radius, $R_{\rm T}$, must be less
than the co-rotation radius, $R_{\rm co}$ (where the Keplerian angular
rotation rate of the disc is the same as the angular rotation rate of
the star), as material outside this radius will have too much angular
momentum to fall onto the star. This is implicitly adopted in the
proposed accretion mechanism, where the disc is magnetically
truncated, and therefore locked to the star. This magnetic truncation
would lead to an inner wall (and inner hole) of the accretion disc.
Evidence for an inner disc wall has been found from near-IR excesses
from the inner rims of Herbig Ae stars, with emission peaking at
2-3$\mu$m \citep{2001ApJ...560..957D}. The actual shape and radial
position of this wall is affected by several factors, discussed
further in Sections \ref{dust_edge} and \ref{inner_edge}. The previous
work of \cite{1997AJ....114..288M} explored the effects of simple mass
accretion and the radius to the inner disc on the infrared colours of
CTTS, using a prescribed range of inner disc radii.
\cite{1997AJ....114..288M} find that the scatter in infrared colours
for there parameter range, $R_{\rm inner}=$1$-$12$R_*$ and
$\dot{M_{\rm acc}}=$10$^{-9}$to 10$^{-5} M_{\odot}yr^{-1}$, matches
the scatter of the CTTS locus found in the data. The models of
\cite{1997AJ....114..288M} used physically flat discs however meaning
their results will not stand for discs with significant flaring. More
importantly for our work \cite{1997AJ....114..288M} find a correlation
of the infrared locus of models with the accretion and inner disc
location. Essentially, changes in the stellar flux levels and position
of the inner wall lead to temperature changes of this inner edge and
therefore changes in the balance of flux across the infrared
wavelength regime. As this study does not include the effects of
vertical hydrostatic equilibrium or even the study of flared discs one
would expect, perhaps, a more complex relationship between the inner
edge position and infrared excess in `real' systems. This means that
studies aiming to provide evidence of disc locking by comparing disc
indicators with rotation rates may include unforeseen biases. As the
rotation rate will effect the co-rotation radius and therefore
truncate the disc at different radii, thereby enhancing or suppressing
any relationship of infrared excess (or disc presence) with rotation
rate.

In summary, several key areas in the theoretical studies of BDs hinge
critically on observational constraints, where the derived values are
heavily affected by disc and/or accretion presence. Formation
mechanisms, planet formation, IMFs, disc dissipation timescales,
accretion rate--stellar mass relationships and studies of disc-locking
mechanisms using correlations of rotation rates with infrared
excesses, all require derivations of either masses, ages or disc
fractions (or a combination), supplemented by other parameters. The
consequences of the disc structure for BD accretion discs,
particularly vertical structure, given the assumption of vertical
hydrostatic equilibrium, have already been discussed for a limited
range of BD parameters in \cite{2004MNRAS.351..607W}. Indeed
\cite{2004MNRAS.351..607W} have shown that BDD systems seen at higher
inclinations can be confused with CTTS systems, having similar colours
and magnitudes. A benchmark model grid of spectra for star and disc
systems in the mass range 0.1$M_{\odot}<M_*<$50$M_{\odot}$ has
previously been published in \cite{2006ApJS..167..256R}.

@@

In this study we assume that accretion from the central star occurs
along magnetically channelled columns from the inner disc boundary.
For TTS stars, \cite{2007prpl.conf..479B} show that the magnetic
truncation radius ($R_{\rm mag}$) is less than the co-rotation radius
($R_{\rm co}$), where the angular Keplerian velocity of the
disc is equal to the surface angular velocity of the central star.
Calculations of the magnetic truncation radius depend on derivations
of the surface magnetic field \citep{1991ApJ...370L..39K}. This is
currently unavailable for BD stars, due to increased molecular species
obscuring the Zeeman splitting signatures, that are normally used to
derive stellar surface magnetic fields. Therefore, for our model grid
we have adopted an initial inner disc radius as the co-rotation
radius,
\begin{equation}
{R_{\rm inner}=({GM_*\tau^2\over{4\pi ^2}})^{1\over{3}}},
\label{inner_eq}
\end{equation}
where $\tau$ is the stellar rotation period and $R_{\rm inner}$ is the
inner radius. This is effectively adopting a disc-locking mechanism
(without associated angular momentum loss), as for disc-locked stars,
$R_{\rm mag}\approx R_{\rm co}$,
\citep{1991ApJ...370L..39K,1994ApJ...429..781S}. Therefore, for our
model simulations, this inner edge radius is dependent on, and derived
from, the adopted value of the rotational period for the central star,
as well as being weakly dependent on the stellar mass. As discussed in
Section \ref{intro}, the inner disc can be cleared through a number of
mechanisms, including binarity or Giant planet formation,
photoevaporation or photoionisation of the disc and dust grain growth
or settling. For BDD systems where a disc is modeled a treatment of
dust sublimation is included (discussed in Section \ref{dust_edge}).
However, the effects of binarity or Giant planet formation are
neglected. Further to the stellar mass and period required prior to
calculation of the inner disc radius, we require a disc mass (in
stellar masses).

@@

\begin{figure}
  \vspace*{174pt}
  \special{psfile="Fig/albedo.ps"
   hoffset=235 voffset=0 hscale=30 vscale=30 angle=90}
 \caption{Figures of the albedo (\textit{top panel}), and scattering
   (dashed line) and absorption (solid line) opacities (\textit{bottom
     panel}) against log($\lambda$) (in $\mu$m) for our adopted dust
   population. For \textit{both panels} the vertical dashed line is
   plotted at 10$\mu$m to highlight the silicate features.
   \label{albedo}}
\end{figure}

@@

The practical implementation is as follows: An initial coarse
approximate temperature structure is found for the grid without dust,
only a fixed gas distribution. Then, first three full \textit{Lucy}
algorithm iterations are run for a grid with a maximum cell opacity of
0.1 (with dust added to this level). Three more iterations are then
performed with dust opacity limited to $\tau=$1 per cell. A further
three \textit{Lucy} algorithm iterations are then performed with the
full dust opacity, where dust is added until the gas-to-dust ratio is
achieved($\epsilon =$100). For each iteration (from the first to the
ninth) dust in cells which reach the sublimation temperature and
density is removed from the disc by setting the opacity to zero for
the cell. Once these nine iterations are complete, the grid is refined
around the disc photosphere, and the radiative transfer algorithm is
iterated through to convergence. After this first convergence, a
vertical hydrostatic equilibrium algorithm is run, after which the
disc grid is unrefined and then re-refined around the disc
photosphere. This process continues for five hydrostatic equilibrium
calculations. Finally, once the hydrostatic equilibrium iterations are
complete we perform a final dust sublimation run to remove any cells
where the dust temperature has moved above the sublimation temperature
and further refine the photosphere and inner edge of the disc (whilst
also unrefining cells where the opacity has reduced significantly and
high resolution is no longer required). In this model the dust
sublimation is effectively digital i.e. when the density and
temperature for sublimation is reached for a given cell the dust is
destroyed. We did attempt a more realistic model where the dust
sublimation causes the dust to be destroyed at an evanescently
decaying rate into the particular region. This was expected to lead to
a more realistic situation where the dust destruction propagates into
the disc material, and does not instantly destroy an entire
cell. However, this implementation lead to incremental sections of
dust being removed after each iteration, resulting in little change in
the overall density structure of the disc. This is perhaps interesting
for real systems, in that the dust sublimation should not be expected
to act as a digital `switch'.

@@

Binarity and Giant planet formation would be expected to clear an
inner disc and shut down accretion from within the inner wall due to
gravitational perturbations. In addition, photoevaporation from UV and
FUV flux is expected to halt accretion from radii within the inner
disc due to the removal of this material via a stellar/disc wind
\citep{2006MNRAS.369..229A}. These two clearing mechanisms could be
differentiated by the resulting masses of the discs
\citep{2007MNRAS.378..369N}. However, for our model systems we have
neglected both of these effects.

@@

Once the radiative transfer code was completed simulated SEDs were
generated. These SEDs can be generated for any distance and for any
system-observer inclination. For our models we have set the distance,
to 10pc to create absolute flux SEDs and selected ten inclinations (0,
27, 39, 48, 56, 64, 71, 77, 84 and 90$^{\circ}$). A further useful
feature of the {\sc torus} code is that the emitted photons which make
up the SED are tagged on their way to the observer. These tags
separate the photons into four groups. Firstly, photons are separated
by source into thermal (disc) or stellar groups. These groups are then
subdivided into photons which reach the observer either directly or
after scattering (for example see Figure \ref{disc_eg_173}). The
resulting SEDs are discussed in Section \ref{seds}.

@@

This section details the range of each of the input parameters we have
varied and, where possible, gives justification for the selected
ranges from published observations. The simulations in this paper
cover variations in the stellar mass, stellar age, stellar rotation
rate, accretion rate, the areal coverage of the accretion stream, disc
mass fraction and the system inclination. In a subsequent paper we
introduce variation of the outer radius and using an analytical
vertical disc structure as oppose to vertical hydrostatic equilibrium.
A summary of the values adopted for each input variable is shown in
Table \ref{par_space_table}.

@@

Typical disc lifetimes for solar type stars are of order 10 Myrs
\citep{2001ApJ...553L.153H}. Therefore, we have adopted input ages of
1 and 10 Myrs for our model grid, to span the approximate range of
ages over which the discs influence will be important.

@@

Little observational evidence can be found for approximate sizes of
accretion hot spots due to their more transient nature and often
smaller coverages, when compared to cooler or `plage' spots
\citep{2007prpl.conf..297H}. \cite{1995A&A...299...89B} modeled the
size of the cool spots on solar-type stars for a selection of
periodically variable candidates. They found projections of cooler
spots, onto the stellar disc, of a few to $~$60\%.
\cite{1995A&A...299...89B} also found projected sizes, onto the
stellar disc, of typically a few \% to around 10\% for hot spots.
\cite{1996AJ....112.2159B} used observations of YY Orionis monitoring
flux amplitude variations as a function of wavelength to derive a
probable hot spot area of around 10\%. The spot temperature was also
modeled for YY Orionis in \cite{1996AJ....112.2159B}, resulting in a
best fitting areal coverage of 11\%. Therefore, to bound the probable
areal coverage range of the accretion hot spots we have adopted areal
coverages of 1 and 10\%.

@@

@@

For our model grid we have selected a range of input parameters
covering expected, from observations, ranges. We have input the
fundamental parameters of age and mass by adoption of a set of
theoretical isochrones. We have intrinsically included a correlation
of rotation rate with initial inner edge position by setting this at
the co-rotation radius, which is expected to result in a correlation
of rotation rate with IR excess \citep{1997AJ....114..288M} (the
reverse of that expected due to angular momentum loss through
disc-locking). We have not assumed a relationship between accretion
rate and stellar mass for our model grid. Therefore, analysis of our
model grid in an attempt to recover the fundamental parameters of age
and mass, and show accretion rate to stellar mass or rotation rate to
infrared excess relationships, will allow us to assess the reliability
of these current theoretical models or relationships.

In this Section we first discuss the physical structure, both density
and temperature, of the BDD disc systems (Section \ref{disc_struct})
across our parameter space. Then we discuss the resulting simulated
SEDs for these BDD systems (Section \ref{sed_analysis}). Finally, we
produce simulated photometric observations (Section \ref{photometry}).
These simulated observations allows us to show that recovery of the
input masses and ages for our accreting BDD systems is unreliable. We
show that despite an intrinsic relationship between rotation rate and
initial inner edge position we do not recover a rotation rate
correlation with IR excess. In addition, despite not intrinsically
including a dependence of accretion rate on stellar mass in our grid.
We show that current observational techniques and theoretical models
applied to the grid would result in a relationship of this type being
derived.
